## Exercise 1
```{r include=FALSE}
library(lme4) 
fruitflies.1 <- read.table("./data/fruitflies.txt", header = TRUE)
```

**a)**
```{r include=FALSE, echo=TRUE}
model_a <- lm(fruitflies.1$longevity ~ fruitflies.1$activity, data = fruitflies.1)
```

When using a mixed effects model using longevity as the dependent variable and activity as the independent variable the following results are obtained:

```{r echo=TRUE}
summary(model_a)
```

When looking at the P-values of the variables $high$ (intercept), $isolated$ and $low$ both these effects are, taking into consideration a signifcance level of 0.05, considered significant. When looking at the estimated longevity we can conclude that the reference category high is 38.720 days. With isolated activity 38.720 + 24.840 = 63.560 days and finally, with low activity 38.720 + 18.040 = 56.760 days.
**b)**


```{r include=FALSE, echo=TRUE}
model_b <- lm(longevity ~ activity * thorax, data = fruitflies.1)
```

When using a mixed effects model using longevity as the dependent variable and activity and thorax as the independent variable the following results are obtained:

```{r echo=TRUE}
summary(model_b)
```
**c)**

**d)**

**e)**
