## Exercise 1
```{r include=FALSE}
library(lme4) 
fruitflies.1 <- read.table("./data/fruitflies.txt", header = TRUE)
fruitflies.1$loglongevity <- log(fruitflies.1$longevity)
```

**a)**

```{r echo=FALSE}
# Load necessary libraries
library(ggplot2)

# Add a column for log-transformed longevity
fruitflies.1$loglongevity <- log(fruitflies.1$longevity)

# Create a boxplot or violin plot
ggplot(fruitflies.1, aes(x = activity, y = longevity)) +
  geom_boxplot() +
  ylim(0, 100) +  # Set y-axis limits
  labs(title = "Effect of Sexual Activity on Longevity of Fruit Flies",
       x = "Group",
       y = "Log(Longevity)")
```

Looking at the boxplots both isolated and low seem to be symmetric, with low having one outlier. High is negatively skewed with one outlier. 

To test if sexual activity influences longlevity, the most obvious model to use is anova. When using this model, we would be presuming the residuals are normally distributed, which are plotted here:

```{r echo=FALSE}
anova_result_a <- aov(loglongevity ~ activity, data = fruitflies.1)
residuals_anova <- residuals(anova_result_a)

par(mfrow = c(1, 2))  # Create a 2x2 grid for plots

hist(residuals_anova, main = "Histogram of Residuals", xlab = "Residuals")  # Histogram
qqnorm(residuals_anova); qqline(residuals_anova)  # Q-Q plot
```

With this data, normality is dubious. So, instead of anova, we decided to use the Krushkal-Wallis, which is a rank-based alternative to anova.

When using a Krushkal-Wallis test, using log-longevity, which is added to the dataframe as an extra column, as the dependent variable and activity as the explanatory variable and the following hypothesizes:

$H_0$: The mean log-longevity is the same for flies with different levels of sexual activity.

$H_a$: At least one group's mean log-longevity is different from the others.

The following results are obtained:


```{r echo=FALSE}
kruskal.test(fruitflies.1$loglongevity, fruitflies.1$activity)
```

Considering a significance level of 0.05 we can conclude that the null hypothesis is rejected and thus that there are significant differences between at least one group's mean log-longevity from the others.

Finally, the following mean longevity for each activity level are obtained:
```{r}
tapply(fruitflies.1$longevity, fruitflies.1$activity, mean)
```
**b)**

```{r echo=FALSE}
anova_result_b <- aov(loglongevity ~ activity + thorax, data = fruitflies.1)
residuals_anova <- residuals(anova_result_b)

par(mfrow = c(1, 2))  # Create a 2x2 grid for plots

hist(residuals_anova, main = "Histogram of Residuals", xlab = "Residuals")  # Histogram
qqnorm(residuals_anova); qqline(residuals_anova)  # Q-Q plot

```
After the Thorax variable is added, normality of residuals seems more presumable, unlike in a.

When using a ANOVA model (Specifically randomized block design) using again log-longevity as the dependent variable and activity and thorax as the explanatory variable and the following hypothesizes:

$H_0$: The mean log-longevity is the same for flies with different levels of sexual activity.

$H_a$: At least one group's mean log-longevity is different from the others.

The following results are obtained:
```{r echo=FALSE}
summary(anova_result_b)
```
Considering a significance level of 0.05 we can conclude that the null hypothesis is rejected for both thorax and activity and thus that there are significant differences between at least one group's mean log-longevity from the others.

```{r echo=FALSE}
# Given coefficients
intercept <- 1.2189341
coef_activity_isolated <- 0.4099810
coef_activity_low <- 0.2857017
coef_thorax <- 2.9789877
```

```{r}
coef(anova_result_b)
```

```{r}
mean_thorax <- mean(fruitflies.1$thorax)

longevity_isolated <- exp(intercept + coef_activity_isolated + coef_thorax * mean_thorax)
longevity_low <- exp(intercept + coef_activity_low + coef_thorax * mean_thorax)
longevity_high <- exp(intercept + coef_thorax * mean_thorax)
```

The mean thorax across the whole population is `r mean_thorax`. Now using the above linear regression we find the estimates are `r longevity_isolated` days for isolated activity, `r longevity_low` days for low activity and `r longevity_high` days for high activity.

**c)**


```{r echo=FALSE}

library(ggplot2)

ggplot(fruitflies.1, aes(x = thorax, y = longevity, color = activity)) +
  geom_point() +
  labs(title = "Scatterplot Across Activity Levels",
       x = "Thorax Length",
       y = "Longevity",
       color = "Activity") +
  facet_wrap(~activity) +  # Separate plots for each activity level
  theme_minimal()

anova_result_b <- aov(longevity ~ activity * thorax, data = fruitflies.1)

```

When using a ANOVA model using longevity as the dependent variable and activity and thorox as the explanatory variable and the following hypothesizes:

$H_0$: The mean longevity is the same for flies with different levels of sexual activity.

$H_a$: At least one group's mean log-longevity is different from the others.

The following results are obtained:
```{r echo=FALSE}
summary(anova_result_b)
```

Considering a significance level of 0.05 we can conclude that the null hypothesis is rejected for both thorax and activity and thus that there are significant differences between at least one group's mean longevity from the others. However, since the combined effect of thorax and activity is above the significance level 0.05 we fail to reject the null hypothesis and thus that the the mean log-longevity is the same for flies with different levels of sexual activity.

In conclusion there seems to be correlation in the scatterplots, but there is no effect on the interaction between activity and thorax.

**d)**


As both thorax and activity are significant as found in earliers question it would be more appropriate to use both values in the selected model. Neither is per defenition wrong. In question a the model purely focusses on activity, b focusses on both sexual activity and thorax length in exploring possible effects.

**e)**

```{r echo=FALSE}
ancova_model <- aov(longevity ~ activity + thorax, data = fruitflies.1)
summary(ancova_model)
```

```{r echo=FALSE}
residuals <- residuals(ancova_model)
par(mfrow = c(1, 2))
# Create a histogram of the residuals
hist(residuals, main = "Histogram of Residuals Log", xlab = "Residuals", ylab = "Frequency", col = "lightblue", border = "black")

residuals <- residuals(anova_result_b)
hist(residuals, main = "Histogram of Residuals non-Log", xlab = "Residuals", ylab = "Frequency", col = "lightblue", border = "black")
```

Considering the fact that for both models significance levels are roughly the same and for both residuals are normally distributed there is no strong reason to use a log conversion. 