## Exercise 1
```{r include=FALSE}
library(lme4) 
fruitflies.1 <- read.table("./data/fruitflies.txt", header = TRUE)

# By default, it's just strings. We want a factor (in C terms: an enum)
fruitflies.1$activity = as.factor(fruitflies.1$activity)

# The assignment asks specifically to use this log. Likely because the relation is not linear.
# It might be smart to also show this with some plots.
# The inverse to log() (which is log e) is exp() (which is e^x) if you want to convert numbers back.
fruitflies.1$loglongevity <- log(fruitflies.1$longevity)
```

**a)**
```{r include=FALSE, echo=TRUE}
# Mixed effect models without any block variables is just one-way anova.
model_a <- lm(fruitflies.1$loglongevity ~ fruitflies.1$activity, data = fruitflies.1)
```

When using one-way anova using longevity as the dependent variable and activity as the independent variable the following results are obtained:

```{r echo=TRUE}
anova(model_a)
```

When looking at the P-values of the variables $high$ (intercept), $isolated$ and $low$ both these effects are, taking into consideration a signifcance level of 0.05, considered significant. When looking at the estimated longevity we can conclude that the reference category high is 38.720 days. With isolated activity 38.720 + 24.840 = 63.560 days and finally, with low activity 38.720 + 18.040 = 56.760 days.
**b)**


```{r include=FALSE, echo=TRUE}
# activity * thorax implies there would be an interaction term. I find this extremely unlikely, as the flies were
# randomly selected to be in each of the groups. (Also, this wasn't a mixed effects model.)
# Random effects also doesn't make a ton of sense, as there is only one block variable.
# Therefore, I think it might be best to use block designs from lecture 5
model_b <- lm(longevity ~ thorax + activity, data = fruitflies.1)
```

When using a randomised block design using longevity as the dependent variable and activity and thorax as the independent variable the following results are obtained:

```{r echo=TRUE}
anova(model_b)
```
**c)**

**d)**

**e)**
