## Exercise 1

**a)** 
```{r include=FALSE}
Ice_cream.1 <- read.csv("./data/Ice_cream-1.csv")
mean_video <- mean(Ice_cream.1$video)
``` 

From the given dataset $icecream.csv$ the mean $\mu$ is `r mean_video`

```{r, echo=FALSE}
# Create a histogram
par(mfrow=c(1,3))
boxplot(Ice_cream.1$video, ylab="Value of Video", main="Boxplot of video")
hist(
  Ice_cream.1$video,
  xlim = c(20, 80),
  xlab="Value of video",
  ylim = c(0, 40),
  main="Histogram of video"
)
qqnorm(Ice_cream.1$video, pch = 1, frame = FALSE, ylim = c(20, 80), main="QQ-plot of video")
qqline(Ice_cream.1$video, col = "steelblue", lwd = 2)
``` 

When analyzing the histogram and QQ-plot above the assumption of a normal distribution can be made as the boxplot looks symmetrical, th histogram is bell shaped and the QQ-plot follows a relatively straight line.

```{r include=FALSE}
Ice_cream.1 <- read.csv("./data/Ice_cream-1.csv")
mean_video <- mean(Ice_cream.1$video)
sd_video <- sd(Ice_cream.1$video)
cl <- 0.97
z_score <- qnorm((1 + cl) / 2)
margin_of_error <- z_score * (sd_video / sqrt(length(Ice_cream.1$video)))

lower_bound <- mean_video - margin_of_error
upper_bound <- mean_video + margin_of_error
``` 
The 97% Confidence Interval for the Mean  $\mu$ is `r mean_video` has a lower bound of `r lower_bound` and an upper bound of `r upper_bound`

**b)** 

**c)** 

**d)** 

**e)** 

**f)** 